Dynamic Adaptive Saturated Sliding Mode Control For Deployment of Tethered Satellite System

### 2016-12-30 manuscript.tex
20161230

In the analysis of the deployment of the TSS, there is a primary assumption that the movement of entire system is similar to a ``rolling ball'' behavior with three DOFs modelled, and it is reasonable to invoke the spherical coordinates to describe the behavior~\cite{Williams2008}. During space mission, the mass of the tether is much smaller than that of the mother satellite, and hence it is always ignored sufficiently. Some other assumptions are obvious and as follows. The sizes of satellites are small compared with the length of the tether, and hence a satellite can be regard as a particle if the tether is long enough. Actually, the mass ratio between the subsatellite and the mother satellite is tiny, i.e., in YES2, the mother satellite mass is 6530~$kg$, and the subsatellite mass is 12~$kg$~\cite{Williams2008}.  Owing to the tiny mass ratio between the satellites, the deployment of the subsatellite will not affect the motion of the mother satellite during the whole mission. The deployment is regulated by the tension and thrusters, and to guarantee the establishment of the ``rolling ball'' model, the tether remains inextensible, straight and stiff all the time.\par

@@

&\bar m l^2\phi''+2\bar m \phi'l'l+\bar m l^2\sin\phi\cos\phi\left[(\theta'+\Omega)^2+3\bar m \Omega^2\cos^2\theta\right]=\tau_\phi+\tau_{\phi d},\label{eq:dynamics 1}

@@

where, $m = m_1+m_2$ stands for the total mass of the system, $m_1$ and $m_2$ of which denote the mass of the mother satellite and of the subsatellite, respectively. $(\cdot)'$ represents a first derivative with respect to time, and $l$ means the tether length connecting the satellites. $\theta$ and $\phi$ depict the roll and pitch dynamics of the TSS. For clarity, $m_1m_2/m$ is rewritten as $\bar m$. $\tau_t$, $\tau_\theta$ and $\tau_\phi$ are the tension, the in-plane thruster force and the out-of-plane thruster force, respectively. $\tau_{t d}$, $\tau_{\theta d}$ and $\tau_{\phi d}$ stand for the bounded external disturbances correspondingly. Typically, the dimensionless transformations simplify issues for researchers to design the deployment scheme, and the conventional transformations are represented as shown below~\cite{williams2009yes2,wen2015space,wen2016constrained}:

@@

Based on the novel dimensionless transformations~(\ref{eq:varcon}), Eq.~(\ref{eq:dynamics 1}) can be converted into the dimensionless Euler-Lagrange dynamical equation described as follows:
\begin{align}
H(\bm q)\ddot {\bm q}+C(\bm q,\dot{\bm q})\dot{\bm q}+G(\bm q) = \bm\tau+\bm\tau_d,\label{eq:dynamic model}
\end{align}
where
\begin{align}
\begin{split}
\bm q&= \begin{pmatrix}\lambda&\theta&\phi\end{pmatrix}^T\\
\bm \tau&=\begin{pmatrix}\hat\tau_t&\hat\tau_\theta&\hat\tau_\phi\end{pmatrix}^T\\
\bm\tau_d&=\begin{pmatrix}\hat\tau_{td}&\hat\tau_{\theta d}&\hat\tau_{\phi d}\end{pmatrix}^T\\
H(\bm q) &= \begin{bmatrix}1 &0&0\\0 &\lambda^2\cos^2\phi&0\\0&0&\lambda^2\end{bmatrix}\\
C(\bm q,\dot{\bm q}) &=\begin{bmatrix}0 &-(\lambda\dot \theta+2\lambda)\cos^2\phi&-\lambda\dot\phi\\(2\lambda+\lambda\dot \theta)\cos^2\phi&\lambda\dot \lambda\cos^2\phi-\lambda^2\dot\phi\sin\phi\cos\phi&-(\dot\theta+2)\lambda^2\sin\phi\cos\phi\\ \lambda\dot\phi&\lambda^2(\dot\theta+2)\sin\phi\cos\phi&\lambda\dot \lambda\end{bmatrix}\\
G(\bm q) &=\begin{bmatrix}-\lambda\cos^2\phi+2\lambda+3\lambda\cos ^2\theta\cos^2\phi\\3\lambda^2\cos\theta\sin\theta\cos^2\phi\\(1+3\cos^2\theta)\lambda^2\sin\phi\cos\phi\end{bmatrix}.
\end{split}
\end{align}
Based on the expression above, one can find that $\dot H(\bm q) - 2C(\bm q,\dot{\bm q})$ is skew-symmetric, namely,
\begin{align}
\bm q^T(\dot H(\bm q) - 2C(\bm q,\dot{\bm q}))\bm q = 0.
\end{align}
\begin{myrem}

@@

In low-altitude orbit , the external disturbance $\bm\tau_d$ contains the gravity gradient torque, the aerodynamic interference and the solar radiation pressure torque. Thanks to existing studies, the values of the solar radiation pressure and the aerodynamic torque are about $1\times10^{-5}$ Nm, and meanwhile the gradient torque is in the range of $1\times10^{-5}$ Nm. The space perturbation can be specifically regarded as the periodic function in this paper, and the period equals the orbital period $5000$ s~\cite{Inamori2015192}. Hence, the external disturbance can be described by the following expression

@@

\bm\tau_d=\left(0.05\sin(2\pi t/5000),0.05\sin(2\pi t/5000),0.05\sin(2\pi t/5000)\right)^T.

@@

@@

where $\varphi_1$ can be selected as a positive scalar $\varphi_1=1$.

@@

\textcolor{red}{In the aforementioned controller design, signum functions in inputs lead to chattering phenomena, which may cause adverse unmodeled dynamics to ruin the control system. Herein, we adopt the boundary layer technique, which has been widely used in engineering applications~\cite{Youndg1999A,Hwang1996sliding,In2000Robust,Chen2005,Yu2009sliding,Jin2009Practical,Al2004sliding}, to reduce chattering through replacing $\rm{sign}\it{(x)}$ by $\rm{sat}\it{(x/\nu)}$, where

@@

where $\hat\bullet$ stands for the measurement of the term $\bullet$ with uncertainties. $\varpi$ is the uncertainty boundary parameter, and here we select $\varpi=2$. $\bm\tau_d$ is the external disturbance meaning  space perturbations, such as gravity gradient torque, aerodynamic interference and solar radiation pressure torque.  In this simulation, the measured terms can be specified as: $\hat{\bar m}=1.1\bar m$ may mean some space debris affixed to the subsatellite; $\bar L=1.02L$ and $\bar \Omega=0.98\Omega$ mean some measurement errors happening to the total length and the orbital angular velocity. In Fig~\ref{fig:6}, although the dynamics have already been  stable, the tension is switching in the high frequency. It is commonplace in the sliding mode control law design, called chattering, since the signum function is used in the input, which may cause unmodeled high frequency dynamics, and for reducing the chattering, we introduce the boundary layer technique into the proposed control laws with the thickness of the boundary layer specified as a positive scalar $\nu=1\times 10^{-4}$ in this simulation~\cite{1039802}.

The simulation results of the deployment dynamics subjected to uncertainties are exhibited from Fig~\ref{fig:7} to Fig~\ref{fig:10}. In Fig~\ref{fig:7}, although there exist parametric uncertainties compared with the dynamics in Figs~\ref{fig:3}-\ref{fig:6}, QST-SMC still performs the deployment with the shortest rise time and the minimum overshoot among the three methods, and the tether is deployed to the desired position at about 8 $rad$ by observing the dynamics of the in-plane and out-of-plane angles from Fig~\ref{fig:9}. In addition, the result in Fig~\ref{fig:10} shows that the tension falls in the reasonable region, and meanwhile the chattering of the input in the stable status is reduced by utilizing the boundary layer technique. \textcolor{red}{The control precisions in Figs~\ref{fig:7}-\ref{fig:9} are almost same to Figs~\ref{fig:3}-\ref{fig:6}, which means the loss of precision caused by the boundary layer technique is limited. Moreover, the simulation results show that the good performance of proposed controller and the suitable boundary layer thickness guarantee that the system state error converges to the small vicinity of origin, which is consistent with the analyses in Theorem~\ref{thm:boundary}.} Hence, the robustness of the proposed control laws can guarantee the stability of the deployment subjected to parametric uncertainties and external disturbances.
\begin{figure}
\centering
\includegraphics[width=0.6\textwidth]{paper4_fig7_20161025.eps}
\caption{Length dynamics of the deployment subjected to uncertainties and disturbances}
\label{fig:7}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.6\textwidth]{paper4_fig8_20161025.eps}
\caption{Length rate dynamics of the deployment subjected to uncertainties and disturbances}
\label{fig:8}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.6\textwidth]{paper4_fig9_20161025.eps}
\caption{Swing dynamics of the deployment subjected to uncertainties and disturbances}
\label{fig:9}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.6\textwidth]{paper4_fig10_20161025.eps}
\caption{Tension produced by utilizing boundary layer technique}
\label{fig:10}
\end{figure}
\section{Conclusion}\label{sec:5}
In this paper, the proposed adaptive sliding mode control laws provide the deployment dynamics of the TSS with a simplified solution for the reaching phase, which ignores the nonlinear nature of the model, and the new dimensionless transformation subjected to the limited input guarantees the effectiveness of the auxiliary Lyapunov function with adaptive terms. The simulation results show that the proposed methods adjust the length of the tether and the swing dynamics as desired, although there exists the input limitation during the deployment. Also, the robustness to parametric uncertainties and external disturbances is verified reasonably. According to the results mentioned above, combining error dynamics design approach, adaptive approach and nonlinear sliding mode surface design technology may provide a further insight into regulation of the deployment dynamics of the TSS.
\section{Acknowledgment}
This work is partially supported by the National Natural Science Foundation of China (No. 61104112, 61673009, 61503101, 91438202).
\section{References}
\bibliography{paper4_ref}
\bibliographystyle{elsarticle-num}
\end{document}

>>>>

In the analysis of the deployment of the TSS, there is a primary assumption that the movement of entire system is similar to a ``rolling ball'' behavior with three DOFs modelled, and it is reasonable to invoke the spherical coordinates to describe the behavior~\cite{Williams2008}. During space mission, the mass of the tether is much smaller than that of the mother satellite, and hence it is always ignored sufficiently. Some other assumptions are obvious and as follows. The sizes of satellites are small compared with the length of the tether, and hence a satellite can be regard as a particle if the tether is long enough. Actually, the mass ratio between the subsatellite and the mother satellite is tiny, i.e., in YES2, the mother satellite mass is 6530~$kg$, and the subsatellite mass is 12~$kg$~\cite{Williams2008}.  Owing to the tiny mass ratio between the satellites, the deployment of the subsatellite will not affect the motion of the mother satellite during the whole mission. The deployment is regulated by the tension and thrusters, and to guarantee the establishment of the ``rolling ball'' model, the tether remains inextensible, straight and stiff all the time. \textcolor{red}{Moreover, for enhancing the control performance of in-plane and out-of-plane dynamics, the thrusters are equipped on the subsatellite to generate thruster forces along the tangent directions of the circle generated by in-plane and out-of-plane angles swing, respectively~\cite{liu2014attitude}}.\par

@@

&\bar m l^2\phi''+2\bar m \phi'l'l+\bar m l^2\sin\phi\cos\phi\left[(\theta'+\Omega)^2+3 \Omega^2\cos^2\theta\right]=\tau_\phi+\tau_{\phi d},\label{eq:dynamics 1}

@@

where, $m = m_1+m_2$ stands for the total mass of the system, $m_1$ and $m_2$ of which denote the mass of the mother satellite and of the subsatellite, respectively. $(\cdot)'$ represents a first derivative with respect to time, and $l$ means the tether length connecting the satellites. $\theta$ and $\phi$ depict the roll and pitch dynamics of the TSS. For clarity, $m_1m_2/m$ is rewritten as $\bar m$. \textcolor{red}{$\tau_t$, $\tau_\theta$ and $\tau_\phi$ are the tension, the in-plane thruster torque and the out-of-plane thruster torque, respectively.} $\tau_{t d}$, $\tau_{\theta d}$ and $\tau_{\phi d}$ stand for the bounded external disturbances correspondingly. Typically, the dimensionless transformations simplify issues for researchers to design the deployment scheme, and the conventional transformations are represented as shown below~\cite{williams2009yes2,wen2015space,wen2016constrained}:

@@

Based on the novel dimensionless transformations~(\ref{eq:varcon}), Eq.~(\ref{eq:dynamics 1}) can be converted into the dimensionless Euler-Lagrange dynamical equation described as follows:
\begin{align}
H(\bm q)\ddot {\bm q}+C(\bm q,\dot{\bm q})\dot{\bm q}+G(\bm q) = \bm\tau+\bm\tau_d,\label{eq:dynamic model}
\end{align}
where
\begin{align}
\begin{split}
\bm q&= \begin{pmatrix}\lambda&\theta&\phi\end{pmatrix}^T\\
\bm \tau&=\begin{pmatrix}\hat\tau_t&\hat\tau_\theta&\hat\tau_\phi\end{pmatrix}^T\\
\bm\tau_d&=\begin{pmatrix}\hat\tau_{td}&\hat\tau_{\theta d}&\hat\tau_{\phi d}\end{pmatrix}^T\\
H(\bm q) &= \begin{bmatrix}1 &0&0\\0 &\lambda^2\cos^2\phi&0\\0&0&\lambda^2\end{bmatrix}\\
C(\bm q,\dot{\bm q}) &=\begin{bmatrix}0 &-(\lambda\dot \theta+2\lambda)\cos^2\phi&-\lambda\dot\phi\\(2\lambda+\lambda\dot \theta)\cos^2\phi&\lambda\dot \lambda\cos^2\phi-\lambda^2\dot\phi\sin\phi\cos\phi&-(\dot\theta+2)\lambda^2\sin\phi\cos\phi\\ \lambda\dot\phi&\lambda^2(\dot\theta+2)\sin\phi\cos\phi&\lambda\dot \lambda\end{bmatrix}\\
G(\bm q) &=\begin{bmatrix}-\lambda\cos^2\phi+2\lambda-3\lambda\cos ^2\theta\cos^2\phi\\3\lambda^2\cos\theta\sin\theta\cos^2\phi\\(1+3\cos^2\theta)\lambda^2\sin\phi\cos\phi\end{bmatrix}.
\end{split}
\end{align}
Based on the expression above, one can find that $\dot H(\bm q) - 2C(\bm q,\dot{\bm q})$ is skew-symmetric, namely,
\begin{align}
\bm q^T(\dot H(\bm q) - 2C(\bm q,\dot{\bm q}))\bm q = 0.
\end{align}
\begin{myrem}

@@

\textcolor{red}{In low-altitude orbit , the external disturbance $\bm\tau_d$ contains the gravity gradient torque, the aerodynamic interference and the solar radiation pressure torque. Thanks to existing studies, the magnitudes of the solar radiation pressure, the aerodynamic torque and the gradient torque are in the range of $1\times10^{-5}$. The space perturbation can be specifically regarded as the periodic function in this paper, and the period equals the orbital period $T_o$ that can be calculated by~\cite{Inamori2015192}
  T_o &= 2\pi/\Omega\\
  \Omega &= \sqrt{\mu_e/R_0^3}\notag
where $\mu_e=3.986\times 10^{14}m^3/s^2$. $R_0$ is related to the orbital altitude and the earth's radius, and in this work, we assume that the tethered satellite system orbits at $260km$ such that $R_0=6.631\times 10^6m$. Using dimensionless transformations~(\ref{eq:varcon}) with $L=3.5km$, the dimensionless disturbance can be described by

@@

\bm\tau_d=\left[1.745\times 10^{-4}\sin(\it{v}),\rm 4.985\times \rm 10^{-8}\sin(\it{v}), \rm 4.985\times 10^{-8}\sin(\it{v})\right]^T.

@@

However, the magnitude of the disturbance is too little to examine the controller performance, hence we intentionally heighten the amplitude in the above expression to obtain the following disturbance formulation, which may express some severe uncertainties or disturbance.
\bm\tau_d=\left[5\times 10^{-2}\sin(\it{v}),\rm 5\times \rm 10^{-2}\sin(\it{v}), \rm 5\times 10^{-2}\sin(\it{v})\right]^T.\notag
where $\it v$ is the dimensionless time.

@@

where $\varphi_1$ can be selected as a positive scalar $\varphi_1=1$.}

@@

In the aforementioned controller design, signum functions in inputs lead to chattering phenomena, which may cause adverse unmodeled dynamics to ruin the control system. Herein, we adopt the boundary layer technique, which has been widely used in engineering applications~\cite{Youndg1999A,Hwang1996sliding,In2000Robust,Chen2005,Yu2009sliding,Jin2009Practical,Al2004sliding}, to reduce chattering through replacing $\rm{sign}\it{(x)}$ by $\rm{sat}\it{(x/\nu)}$, where

@@

where $\hat\bullet$ stands for the measurement of the term $\bullet$ with uncertainties. $\varpi$ is the uncertainty boundary parameter, and here we select $\varpi=2$. $\bm\tau_d$ is the external disturbance meaning  space perturbations, such as gravity gradient effect, aerodynamic interference and solar radiation pressure torque.  In this simulation, the measured terms can be specified as: $\hat{\bar m}=1.1\bar m$ may mean some space debris affixed to the subsatellite; $\bar L=1.02L$ and $\bar \Omega=0.98\Omega$ mean some measurement errors happening to the total length and the orbital angular velocity. In Fig~\ref{fig:6}, although the dynamics have already been  stable, the tension is switching in the high frequency. It is commonplace in the sliding mode control law design, called chattering, since the signum function is used in the input, which may cause unmodeled high frequency dynamics, and for reducing the chattering, we introduce the boundary layer technique into the proposed control laws with the thickness of the boundary layer specified as a positive scalar $\nu=1\times 10^{-4}$ in this simulation~\cite{1039802}.

The simulation results of the deployment dynamics subjected to uncertainties are exhibited from Fig~\ref{fig:7} to Fig~\ref{fig:10}. In Fig~\ref{fig:7}, although there exist parametric uncertainties compared with the dynamics in Figs~\ref{fig:3}-\ref{fig:6}, QST-SMC still performs the deployment with the shortest rise time and the minimum overshoot among the three methods, and the tether is deployed to the desired position at about 8 $rad$ by observing the dynamics of the in-plane and out-of-plane angles from Fig~\ref{fig:9}. In addition, the result in Fig~\ref{fig:10} shows that the tension falls in the reasonable region, and meanwhile the chattering of the input in the stable status is reduced by utilizing the boundary layer technique. \textcolor{red}{Fig.~\ref{fig:11} shows the in-plane and out-of-plane thruster forces. From the results, we can find that all the forces are reasonable and limited in the boundary, and have less chattering by using the boundary layer technique. The in-plane thruster force has the similar results like the tension, the chattering of which is almost eliminated in the final stage. The boundary layer doesn't work well on the out-of-plane thruster force, and it is easy to see that the crossing boundary layer happens a lot, which may be caused by the external disturbance and uncertainties. However the technique still represses the chattering to some extents. Hence we can say that the boundary layer technique can indeed improve the inputs and repress the chattering.} The control precisions in Figs~\ref{fig:7}-\ref{fig:9} are almost same to Figs~\ref{fig:3}-\ref{fig:6}, which means the loss of precision caused by the boundary layer technique is limited. Moreover, the simulation results show that the good performance of proposed controller and the suitable boundary layer thickness guarantee that the system state error converges to the small vicinity of origin, which is consistent with the analyses in Theorem~\ref{thm:boundary}. Hence, the robustness of the proposed control laws can guarantee the stability of the deployment subjected to parametric uncertainties and external disturbances.
\includegraphics[width=0.6\textwidth]{paper4_fig7_20161025.eps}
\caption{Length dynamics of the deployment subjected to uncertainties and disturbances}
\label{fig:7}
\includegraphics[width=0.6\textwidth]{paper4_fig8_20161025.eps}
\caption{Length rate dynamics of the deployment subjected to uncertainties and disturbances}
\label{fig:8}
\includegraphics[width=0.6\textwidth]{paper4_fig9_20161025.eps}
\caption{Swing dynamics of the deployment subjected to uncertainties and disturbances}
\label{fig:9}
\includegraphics[width=0.6\textwidth]{paper4_fig10_20161025.eps}
\caption{Tension produced by utilizing boundary layer technique}
\label{fig:10}
\includegraphics[width=0.6\textwidth]{paper4_response_20161230.eps}
\caption{Thruster force}
\label{fig:11}
\section{Conclusion}\label{sec:5}
In this paper, the proposed adaptive sliding mode control laws provide the deployment dynamics of the TSS with a simplified solution for the reaching phase, which ignores the nonlinear nature of the model, and the new dimensionless transformation subjected to the limited input guarantees the effectiveness of the auxiliary Lyapunov function with adaptive terms. The simulation results show that the proposed methods adjust the length of the tether and the swing dynamics as desired, although there exists the input limitation during the deployment. Also, the robustness to parametric uncertainties and external disturbances is verified reasonably. According to the results mentioned above, combining error dynamics design approach, adaptive approach and nonlinear sliding mode surface design technology may provide a further insight into regulation of the deployment dynamics of the TSS.
\section{Acknowledgment}
This work is partially supported by the National Natural Science Foundation of China (No. 61104112, 61673009, 61503101, 91438202).
\section{References}
\bibliography{paper4_ref}
\bibliographystyle{elsarticle-num}
\end{document}
